Intrusion Detection with Genetic Algorithms and Fuzzy Logic

### 2013-10-13 paper.tex
Started writing the section on the genetic algorithm implementation.

\keywords{Intrusion detection, genetic algorithm, fuzzy logic}

@@

\section{Fuzzy Genetic Algorithm Implementation}
The focus of the research is on detecting new or unknown types of attacks in a network. The intrusion detection system is able to identify normal network activity as well as classify different attack types using a fuzzy genetic algorithm. This kind of algorithm is able to learn new attacks, and has a high detection rate. The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.

@@

\mycomment{Ask Nic how to put in algorithm.}

@@

\mycomment{Ask Nic how to put in algorithm.}

@@

\mycomment{Explain the 8 features}

@@

\mycomment{Ask Nic how to put in algorithm.}

@@

\mycomment{Fix table}

@@

\subsection{Algorithm Overview}

>>>>

\keywords{Intrusion detection, genetic algorithm, fuzzy logic, KDD99, RLD09}

@@

\section{Fuzzy Genetic Algorithm Implementation}
The focus of the research is on detecting new or unknown types of attacks in a network. The intrusion detection system is able to identify normal network activity as well as classify different attack types using a fuzzy genetic algorithm. This kind of algorithm is able to learn new attacks, and has a high detection rate. The system used is evaluated in terms of detection speed, detection rate, and false alarm rate.~\cite{6496342, 6559603}

@@

\mycomment{Figure out how to put in algorithm.}

for each record

   for each rule

      for each attribute

         prob = fuzzy();

         totalprob = totalprob + prob;

      If (totalprob > threshold)

         class is attack;

      else

         class is normal;

      compare the predicted result with actual result

\mycomment{put in alpha and beta symbols}

      find A, B, alpha, and beta,

   calculate fitness

   // create next generation

   preserve\_best()

   crossover()

   mutation()

// A is total number of attack records. B is total number of normal records. alpha is total number of attack records correctly identified as attack. beta is total number of normal records incorrectly classified as attack (false positive).

@@

\mycomment{Figure out how to put in algorithm.}

if (data value is between b and c)

   prob =1.0

else if (data value is between a and b)
\begin{equation*}
   prob = \frac{data - a}{b-a}
\end{equation*}

else if (data value is between c and d)
\begin{equation*}
   prob = \frac{d-data}{d-c}
\end{equation*}

else

   prob = 0.0

@@

duration is the length of the connection in seconds. src\_bytes is the number of bytes sent from source to destination. num\_failed\_logins is the number of failed login attempts. root\_shell returns 1 if root shell is obtained and 0 otherwise. num\_access\_files is the number of operations on access control files. srv\_count is the number of connections to the same service as the current connection in the past two seconds. serror\_rate is the percentage of connections that have ``SYN" errors.
\mycomment{What are SYN errors?}
 same\_srv\_rate is the percentage of connections to the same service.~\cite{KDD99Features}

@@

\mycomment{Figure out how to put in algorithm.}

If (dos\_rule = yes or probe\_rule = yes)

   This record is an attack;

else

   This record is normal;

@@

\mycomment{Fix tables}

@@

\subsection{Algorithm Overview}
The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used a genetic algorithm to make their intrusion detection system. The system is divided into two phases: a precalculation phase and a detection phase. In the precalculation phase, a set of chromosomes are created using training data. Then this set of chromosomes is used in the detection phase for comparison. The following algorithm is used in the precalculation phase:

\mycomment{Figure out how to put in algorithm.}

Algorithm : Initialize chromosomes for comparison

Input : Network audit data (for training)

Output : A set of chromosomes

Range = 0.125

For each training data

   If it has neighboring chromosome within Range

      Merge it with the nearest chromosome

   Else

      Create new chromosome with it

   End if

End for

In the detection phase, a population is created for test data and then the type of the data is predicted. The set of chromosomes that was created in the precalculation phase is used in the detection phase to find the fitness of each chromosome in the population. The following algorithm is used in the detection phase:

\mycomment{Figure out how to put in algorithm.}

Algorithm : Predict data/intrusion type (using GA)

Input : Network audit data (for testing), Precalculated set of chromosomes

Output : Type of data.

Initialize the population

CrossoverRate = 0.15, MutationRate = 0.35

While number of generation is not reached

   For each chromosome in the population

      For each precalculated chromosome

         Find fitness

      End for

      Assign optimal fitness as the fitness of that chromosome

   End for

   Remove some chromosomes with worse fitness

   Apply crossover to the selected pair of chromosomes of the population

   Apply mutation to each chromosome of the population

End while

The authors of~\cite{DBLP:journals/corr/abs-1204-1336} used the KDD99 data set. For the training data set, the KDD99 10\% version file was used, and for the testing data set, the KDD99 corrected version file was used.
